# tpiscaleR

## Terminology 

#### TPI:
The topographic position index is a scale-dependent terrain index. Positiv values indicate a convex and therefore exposed position. 
Negativ values indicate a concave position. Because of its scale dependency, its hard to determine the right radius for the TPI. 
In most cases, the TPI will be used to model another environmental variable like soil moisture.

#### Radius: 
The scale at which the TPI is computed. A radius of 5 e.g. compares the center pixel with the pixels at an 5m radius. 

#### Target Raster:
A one-layered SpatRaster showing the target variable, e.g. soil moisture, NDVI, snow depth. 

#### DSM:
The digital elevation model thats used to calculate the TPI.

## Key features

With this package its possible:

- to evaluate the correlation of TPI at an specific radius to the target raster. Instead of computing the TPI for the whole DSM,
it only uses a certain number of points to evaluate the correlation. This can safe a lot of time. The function is called tpi_sample.

- to find the "best radius" with the highest correlation to the target raster using a bayesian optimaziation. The function is called tpi_opt.

## Requirements

- terra package (to handle the raster data)
- rBayesianOptimization

## Example 1 (testing one specific scale / tpi_sample)

Given we want to research the snowdepth and its relationship to the TPI and we want to check out 
the spearman correlation of a given TPI radius (15 meters). In this case we use
50 pixels to estimate the correlation. 

```{r}


dem <- terra::rast("C:/Users/miles/OneDrive/Dokumente/danield端sentrieb/dsm_snowfree.tif")
snowdepth <- terra::rast("C:/Users/miles/OneDrive/Dokumente/danield端sentrieb/snowdepth.tif")

spearman_15m <- tpiscaleR::tpi_sample(dem, snowdepth, 50, 15, relationship = "spearman")

print(spearman_15m$Score)
```

## Example 2 (Optimizing the tpi radius using bayesianOptimatziation)

Given we want to research the snowdepth and its relationship to the TPI and we want to find out 
the optimalTPI radius. Because we expect a non linear relationship we pick the RSME of a linear regression
as the to be optimized correlation coefficient. We try to find out the optimal radius between 5 and 25 meters using 
a sample of 55 pixels for each radius. We use 5 bayesian iterations and a kappa of 3. Because the bayesian optimization only look for the highest value, the rmse is always negative .
```{r}

dem <- terra::rast("C:/Users/miles/OneDrive/Dokumente/danield端sentrieb/dsm_snowfree.tif")
snowdepth <- terra::rast("C:/Users/miles/OneDrive/Dokumente/danield端sentrieb/snowdepth.tif")

optimal_radius <- tpiscaleR::tpi_opt(5, 25, dem, snowdepth, 55, 5, 3, correlation_coefficient = "rmse_linear")

summary_df <- optimal_radius$History
plot(summary_df$radius, summary_df$Value)

```

 
